---
title: "Defining dock layouts"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Defining dock layouts}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

```{r setup}
library(blockr.core)
library(blockr.dock)
```

A `dock_board` exposes a single argument that controls panel arrangement:
the `layouts`. This vignette walks through every shape `layouts` accepts,
shows what UI each shape produces, and ends with patterns for the layouts
you'll build most often.

# The big picture

Internally, every board carries a `dock_layouts` collection. This is a
named list of one or more **views** (the global tabs you see at the top
of the app). A view's content is a `dock_layout`: the per-view DockView
grid that arranges block and extension **panels**.

Single-page boards are the degenerate case: a `dock_layouts` with one
auto-named view called `"Page"`. The view-nav dropdown is always present;
it just has one entry until you add more.

When you pass a `layouts =` argument to `new_dock_board()`, the constructor
normalises it to a `dock_layouts`:

| You pass | Treated as | Final shape |
| --- | --- | --- |
| A named `list(A = ..., B = ...)` | multi-view (each name is a view) | `dock_layouts` collection |
| A `dock_layout` (already resolved) | single page | wrapped as `Page` slot |
| A `dock_grid` (unresolved spec) | single-page, resolved on the way in | wrapped as `Page` slot |
| A raw list (grid spec) | single-page raw grid, resolved on the way in | wrapped as `Page` slot |

```{r coercion-flow, eval=TRUE, echo=FALSE}
blockr.core::include_mermaid("coercion-flow")
```

The dispatch is name-driven: a list with top-level names becomes
multi-view (one slot per name); an unnamed list (or a `dock_grid`) is
treated as a single-page grid spec. To get multiple views, name the
top-level entries.

So you only ever need to think about two things: the **grid syntax** for
arranging panels inside a view, and **named-list syntax** for having
more than one view.

# Starting from scratch

The simplest possible board: no blocks, no extensions, no `layouts`. Since `new_dock_board()` defaults to `layouts = default_grid(blocks, extensions)` — which returns an empty `dock_grid` here — and the empty grid gets resolved and wrapped as a single `Page` view, the resulting collection is a `dock_layouts` with one empty `Page` slot:

```{r}
new_dock_board()
```

```
┌────────────────────────────┐
│  Page  [+]                 │
├────────────────────────────┤
│                            │
│             ⊕              │
│                            │
│   Start by adding a panel  │
│                            │
│       [ Add panel ]        │
│                            │
└────────────────────────────┘
```

You get a single auto-named `Page` view holding the watermark prompt.
Clicking **Add panel** would normally open the panel picker, but here
the board has no blocks or extensions to choose from, so the picker
says so and points you to add a block first. From this empty starting
point you can grow the board interactively.

When you do supply blocks or extensions but still no `layouts`, the
default `Page` is auto-populated by `default_grid()`: blocks alone
become a single row of panels; blocks together with extensions become
the familiar sidebar-and-main shape (extensions on the left, blocks on
the right). Pass `layouts = ` only to override that.

# Single-page raw grid

The simplest layout: a list (or character vector) of block/extension
IDs. The shape of the list determines the grid.

The two rules to internalise:

1. **List nesting alternates orientation.** The top level lays its
   children out horizontally; one level of nesting flips to vertical;
   another level flips back to horizontal, and so on.
2. **Character vectors create tabs.** A vector of IDs lives inside one
   DockView panel; a list of IDs gets split into multiple panels.

## One panel

A single ID gives one panel filling the whole view:

```{r}
new_dock_board(
  blocks = c(a = new_dataset_block()),
  layouts = list("a")
)
```

```
┌────────────────────────────┐
│  Page  [+]                 │
├────────────────────────────┤
│                            │
│            a               │
│                            │
└────────────────────────────┘
```

## Two panels side by side

Two top-level entries → two columns split horizontally:

```{r}
new_dock_board(
  blocks = c(a = new_dataset_block(), b = new_head_block()),
  layouts = list("a", "b")
)
```

```
┌────────────────────────────┐
│  Page  [+]                 │
├─────────────┬──────────────┤
│             │              │
│      a      │      b       │
│             │              │
└─────────────┴──────────────┘
```

## Two panels stacked vertically

Wrap the entries in **one extra layer** of `list()` to introduce a
vertical split:

```{r}
new_dock_board(
  blocks = c(a = new_dataset_block(), b = new_head_block()),
  layouts = list(list("a", "b"))
)
```

```
┌────────────────────────────┐
│  Page  [+]                 │
├────────────────────────────┤
│            a               │
├────────────────────────────┤
│            b               │
└────────────────────────────┘
```

The outer list still describes a horizontal split, but with only one
child that "split" is a single full-width column. The inner `list("a",
"b")` is at depth 1, so it splits **vertically**: `a` stacks on top of
`b`.

## Tabs (multiple views in one panel)

Use a **character vector** (not a list) to put multiple panels in the
same DockView panel as tabs:

```{r}
new_dock_board(
  blocks = c(a = new_dataset_block(), b = new_head_block()),
  layouts = list(c("a", "b"))
)
```

```
┌────────────────────────────┐
│  Page  [+]                 │
├────────────────────────────┤
│ ┌──────┐┌──────┐           │
│ │  a   ││  b   │           │
│ └──────┘└──────┘           │
│                            │
│  (a is shown, b is a tab)  │
│                            │
└────────────────────────────┘
```

`list("a", "b")` (panels split) and `list(c("a", "b"))` (panels
tabbed) look almost identical in source but produce very different
UIs. The list/vector distinction flips between *split a panel* and
*tabify a panel*.

# Nested grids

Combine the two rules to build any layout.

## Two columns, the right one stacked

```{r}
new_dock_board(
  blocks = c(
    a = new_dataset_block(),
    b = new_head_block(),
    c = new_head_block()
  ),
  layouts = list("a", list("b", "c"))
)
```

```
┌────────────────────────────┐
│  Page  [+]                 │
├─────────────┬──────────────┤
│             │      b       │
│      a      ├──────────────┤
│             │      c       │
└─────────────┴──────────────┘
```

## Two columns, both stacked

```{r}
new_dock_board(
  blocks = c(
    a = new_dataset_block(),
    b = new_head_block(),
    c = new_head_block(),
    d = new_head_block()
  ),
  layouts = list(list("a", "b"), list("c", "d"))
)
```

```
┌────────────────────────────┐
│  Page  [+]                 │
├─────────────┬──────────────┤
│      a      │      c       │
├─────────────┼──────────────┤
│      b      │      d       │
└─────────────┴──────────────┘
```

## Three rows in one column

Add a third level of nesting to flip back to horizontal inside the
vertical stack. Useful when you want a row that holds two panels
side-by-side:

```{r}
new_dock_board(
  blocks = c(
    a = new_dataset_block(),
    b = new_head_block(),
    c = new_head_block()
  ),
  layouts = list(list("a", list("b", "c")))
)
```

```
┌────────────────────────────┐
│  Page  [+]                 │
├────────────────────────────┤
│            a               │
├─────────────┬──────────────┤
│      b      │      c       │
└─────────────┴──────────────┘
```

## Sidebar + tabs

A common dashboard shape: a narrow column on the left holding an
extension, and a tabbed panel on the right with several blocks:

```{r}
new_dock_board(
  blocks = c(a = new_dataset_block(), b = new_head_block()),
  extensions = new_edit_board_extension(),
  layouts = list("edit_board_extension", c("a", "b"))
)
```

```
┌────────────────────────────┐
│  Page  [+]                 │
├─────────┬──────────────────┤
│         │ ┌────┐┌────┐     │
│  edit   │ │ a  ││ b  │     │
│         │ └────┘└────┘     │
│         │                  │
│         │   (tabs)         │
└─────────┴──────────────────┘
```

This is also what `default_grid()` produces when both blocks and
extensions are present: extensions on the left, blocks on the right.

# Multiple views (pages)

Pass a named list. Each named entry becomes a separate page in the
view-nav dropdown:

```{r}
new_dock_board(
  blocks = c(a = new_dataset_block(), b = new_head_block()),
  extensions = new_edit_board_extension(),
  layouts = list(
    Analysis = list("a", "b"),
    Editor = list("edit_board_extension")
  )
)
```

`Analysis` view (active by default, since the first view wins):

```
┌────────────────────────────┐
│ Analysis  [+]   ← view nav │
├────────────────────────────┤
│      a       │      b      │
└────────────────────────────┘
```

Switching to `Editor` via the dropdown:

```
┌────────────────────────────┐
│ Editor  [+]                │
├────────────────────────────┤
│                            │
│           edit             │
│                            │
└────────────────────────────┘
```

Each slot value follows exactly the same grid syntax as the single-page
form: character vectors for tabs, nested lists for splits.

## Choosing the initially active view

By default, the first layout is active. To start on a different one,
construct that view's content with `dock_grid(..., active = TRUE)`:

```{r}
new_dock_board(
  blocks = c(a = new_dataset_block(), b = new_head_block()),
  extensions = new_edit_board_extension(),
  layouts = list(
    Analysis = list("a", "b"),
    Editor = dock_grid("edit_board_extension", active = TRUE)
  )
)
```

```
┌────────────────────────────┐
│ Editor  [+]   ← starts here│
├────────────────────────────┤
│           edit             │
└────────────────────────────┘
```

`dock_grid(..., active = TRUE)` attaches an `active` attribute to
the layout spec. The validator rejects more than one layout marked
`active`. If none is marked, the first one wins.

## Empty views

A view with no panels is fine. Switching to it shows the same
watermark prompt as the empty default board (see `Starting from
scratch`), scoped to that tab.

```{r}
new_dock_board(
  blocks = c(a = new_dataset_block(), b = new_head_block()),
  layouts = list(
    Analysis = list("a", "b"),
    Empty = list()
  )
)
```

# Putting it all together

A pot-pourri exercising every feature: multiple views, nested grids,
tabbed panels, an extension as a sidebar, and an explicit active view.

```{r}
new_dock_board(
  blocks = c(
    raw = new_dataset_block(),
    cleaned = new_head_block(),
    summary = new_head_block(),
    plot1 = new_scatter_block(),
    plot2 = new_scatter_block()
  ),
  extensions = new_edit_board_extension(),
  links = list(
    new_link("raw", "cleaned", "data"),
    new_link("cleaned", "summary", "data"),
    new_link("cleaned", "plot1", "data"),
    new_link("cleaned", "plot2", "data")
  ),
  layouts = list(
    Data = list("edit_board_extension", c("raw", "cleaned")),
    Analysis = list(list("summary", "plot1"), "plot2"),
    Charts = dock_grid(c("plot1", "plot2"), active = TRUE)
  )
)
```

The board has three views. `Charts` is marked active, so the user lands
there first.

**Charts** (active on load): one tabbed panel with `plot1` shown and
`plot2` selectable as a tab.

```
┌────────────────────────────┐
│  Charts  [+]               │
├────────────────────────────┤
│ ┌───────┐┌───────┐         │
│ │ plot1 ││ plot2 │         │
│ └───────┘└───────┘         │
│                            │
│  (plot1 shown, plot2 tab)  │
│                            │
└────────────────────────────┘
```

**Data**: extension on the left, two dataset blocks tabbed on the
right. Same pattern as the "Sidebar + tabs" example earlier, just
inside a named view.

```
┌────────────────────────────┐
│  Data  [+]                 │
├─────────┬──────────────────┤
│         │ ┌─────┐┌────────┐│
│  edit   │ │ raw ││cleaned ││
│         │ └─────┘└────────┘│
│         │                  │
│         │  (data tabs)     │
└─────────┴──────────────────┘
```

**Analysis**: two top-level columns. The left column is a vertical
stack (`list("summary", "plot1")`) and the right column is a single
panel (`"plot2"`).

```
┌────────────────────────────┐
│  Analysis  [+]             │
├─────────────┬──────────────┤
│   summary   │              │
├─────────────┤    plot2     │
│    plot1    │              │
└─────────────┴──────────────┘
```

# Cheat-sheet

| Goal | Syntax |
| --- | --- |
| One panel | `list("a")` |
| Two side-by-side panels | `list("a", "b")` |
| Two stacked panels | `list(list("a", "b"))` |
| Tabbed single panel | `list(c("a", "b"))` |
| Sidebar + main | `list("ext", "main")` |
| Two columns, both stacked | `list(list("a", "b"), list("c", "d"))` |
| Multiple views | `list(A = ..., B = ...)` |
| Start on view `B` | `list(A = ..., B = dock_grid(..., active = TRUE))` |
| Empty starter view | `list(Page = list())` |

# Where to go from here

- See `?dock_grid` and `?dock_layout` for the per-view spec and resolved
  forms; `?active_view` for switching the active view at runtime.
